---
title: "intensity_test"
author: "Magdalena"
date: "2026-05-14"
output: html_document
---

```{r setup, include=FALSE}
#install.packages("sjPlot")
library(tidyverse)
library(emuR)
library(wrassp)
library(ggplot2)
#library(MASS)
library(dplyr)
#library(rms)
library(stringr)
library(performance)
library(car)
library(ordinal)
library(emmeans)
library(sjPlot)
```

```{r}
targetDir <- "C:/Users/magdi/OneDrive/Desktop/ipsR/emu_databases"
```
#load emuDB
```{r}
DB <- load_emuDB(file.path(targetDir, "all_emuDB"),
                       verbose = FALSE)
list_ssffTrackDefinitions(DB)
```

#add track definition for intensity 
#this only has to be executed once 
```{r}
#remove_ssffTrackDefinition(DB, "intensity", deleteFiles = TRUE)
add_ssffTrackDefinition(DB,
                    "intensity",
                     onTheFlyFunctionName = "rmsana")
#list_ssffTrackDefinitions(DB)
```

#list formals of rmsana
```{r}
formals("rmsana")
```

#query all consonants from DB
```{r}
cons_all = query(DB, "dur = C")
cons_all<- requery_hier(DB, 
                    seglist = cons_all, 
                    level = "voice")
cons_all<- cons_all %>% filter(labels!="EL")
cons_all<-requery_hier(DB, 
                    seglist = cons_all, 
                    level = "phon")
cons_int = get_trackdata(DB, cons_all, "intensity")
view(cons_int)
```

#exclude deleted consonants (elision)
```{r}
cons = query(DB, "dur = C")
cons<- requery_hier(DB, 
                    seglist = cons, 
                    level = "voice")
cons<- cons %>% filter(labels!="EL")

```

#cleanup
```{r}
#this was cleanup
cons$labels <- str_replace(cons$labels, "[?]", "")
cons$labels <- str_replace(cons$labels, "VFRIC", "VDFRIC")
cons$labels <- str_replace(cons$labels, "VDRFIC", "VDFRIC")
cons$labels <- str_replace(cons$labels, "VDRIC", "VDFRIC")
cons$labels <- str_replace(cons$labels, "VCLO", "VDCLO")
cons$labels <- str_replace(cons$labels, "REL(.)", "REL")
#all_param <- all_param[-c(152),]
view(all_param)
```

#add column for voice
```{r}

#add column for voice (has 3 categories (voiced/partially voiced/voiceless))
cons$Voice <- ifelse(cons$labels == "VDFRIC" |cons$labels == "APPR"| cons$labels == "VDCLO->REL", "voiced",
                         ifelse(cons$labels == "VLFRIC" |cons$labels == "VLCLO->REL", "voiceless", "partiallyvoiced"))
#view(cons)
sum_voice<- cons |> group_by(Voice, ) |> summarise(n=n())
view(sum_voice)

```

#query v2
```{r}
v2 = query(DB, "dur = V2 | nV2")
v2<-requery_hier(DB, 
                    seglist = v2, 
                    level = "voice")
v2_rms <- get_trackdata(DB, v2, "intensity")

view(v2_rms)
```

#query v1
```{r}
v1 = query(DB, "dur = V1|nV1")
v1<-requery_hier(DB, 
                    seglist = v1, 
                    level = "voice")
v1_rms = get_trackdata(DB, v1, "intensity")
##view(v1_rms)
```

#select minimum int value for cons (in order to calculate ratio (hch is vmax-cmin))
```{r}
min_cons<- cons_int%>% group_by(bundle, labels) %>%
  summarise(T1 = min(T1))%>%
  rename(T1_c=T1)%>%
  ungroup()
##view(min_cons)
```

#select max int value for v2
```{r}
max_v2<- v2_rms %>% group_by(bundle, labels) %>%
  summarise(T1 = max(T1))%>%
  rename(T1_v=T1)%>%
  ungroup()
##view(max_v2)
```

#select max int value for v1
```{r}
max_v1<- v1_rms %>% group_by(bundle, labels) %>%
  summarise(T1 = max(T1))%>%
  rename(T1_v=T1)%>%
  ungroup()
##view(max_v1)
```

#merge min_cons and max_v2
```{r}
int_ratio <- merge(min_cons, max_v2, by="bundle")
int_ratio<- int_ratio %>%
  rename(phon=labels.x, v2=labels.y, T1_v2=T1_v)
##view(int_ratio)
```

#merge v1
```{r}
int_ratio <- merge(int_ratio, max_v1[c("T1_v", "bundle")], by="bundle")
int_ratio<- int_ratio %>%
 rename(T1_v1=T1_v)
##view(int_ratio)
```

# calculate VCR (both with v2 (VCR) and v1 (VCR_v1))
```{r}
int_ratio<- int_ratio%>% mutate(VCR=T1_v2 - T1_c)
int_ratio<- int_ratio%>% mutate(VCR_v1=T1_v1 - T1_c)
#int_ratio<- int_ratio[order(int_ratio$VCR),]
#view(int_ratio)

```

```{r}
int_ratio_voice <- merge(int_ratio, cons, by="bundle")
#view(int_ratio_voice)
sum_voice<- int_ratio_voice |> group_by(Voice, phon) |> summarise(n=n())
#view(sum_voice)
```

#trim white space
```{r}
int_ratio$phon <- trimws(int_ratio$phon)
```

#plot relationship between VCR and VCR_v1 to test if they corelate (the should and do)--> both are probably fine?
# maybe v2 is slightly better, because there are less liquides after than preceeding
```{r}
ggplot(int_ratio, aes(x=VCR, y=VCR_v1)) + geom_point() + geom_smooth() +geom_smooth(method = "loess",
              formula = y ~ x, col="#9966FF") + theme_bw()+ labs(x= "VCR (V2)", y="VCR (V1)")
```
```{r}
ggplot(all_param, aes(x=T1_v2, y=T1_v1)) + geom_point() + geom_smooth() +geom_smooth(method = "loess",
              formula = y ~ x, col="#9966FF") + theme_bw()+ labs(x= "Intensität V2", y="Intensität V1")
cor.test(all_param$T1_v1, all_param$T1_v2, alternative = "greater")
```

#violin plot and boxplot of rel dur per phon
```{r}
ggplot(all_param, aes(x = phon.x, y = dur_rel)) +
  geom_violin( color="#339966") +
  labs(x="Realisierung") +
  geom_boxplot(width = 0.2, fill = "white", color = "#9966FF")+
  theme_bw()
```

#violin plot and boxplot of VCR per phon
```{r}
ggplot(int_ratio, aes(x = phon, y = VCR)) +
  geom_violin( color="#339966") +
  labs(x="Realisierung") +
  geom_boxplot(width = 0.2, fill = "white", color = "#9966FF")+
  theme_bw()
```

# v2 slightly lower in all measures
```{r}
summary(int_ratio$VCR)
summary(int_ratio$VCR_v1)
```

# load rel_dur and V_C_ratio csvs
```{r}
rel_dur<- read.csv("C:/Users/magdi/OneDrive/Desktop/ipsR/data/rel_dur.csv")
rel_dur$phon <- trimws(rel_dur$phon)
#view(rel_dur)
VC_ratio<- read.csv("C:/Users/magdi/OneDrive/Desktop/ipsR/data/VC_ratio.csv")
VC_ratio$phon <- trimws(VC_ratio$phon)
#view(VC_ratio)
```

#merge them all together
```{r}
all_param <- merge(int_ratio, VC_ratio, by="bundle")
all_param <- merge(all_param, rel_dur[c("bundle", "dur_rel")], by="bundle")
#all_param <- all_param %>% 
  #select("bundle", "phon.x", "phon.y", "T1_c", "T1_v", "VCR", "VC_duration", "V_C_ratio", "word", "dur_rel")
all_param <-merge(all_param, cons[c("bundle", "Voice")], by="bundle")
all_param$phon.x <- factor(all_param$phon.x, levels=c( "β̞", "β", "b") ,ordered = TRUE)
levels(all_param$phon.x)
all_param$v2 <- factor(all_param$v2)
#levels(all_param$v2)
all_param$Voice <- factor(all_param$Voice, levels=c("voiced", "partiallyvoiced", "voiceless"), ordered = TRUE)
view(all_param)
```

#plot VCR as violinplot by speaker
```{r}
ggplot(all_param, aes(x = phon.x, y = VCR)) +
  geom_violin( color="#339966") +
  labs(x="Realisierung") +
  facet_wrap(~session, labeller = labeller(session = 
    c("_0283_TUX_jungII" = "283: TUX, jung, m",
      "_0356_PASS_jungII" = "356: PASS, jung, w",
      "_0541_GAWE_jungII" = "541: GAWE, jung, m")))+
  geom_boxplot(width = 0.2, fill = "white", color = "#9966FF")+
  theme_bw()
```

# boxplot with jitter
```{r}
ggplot(all_param, aes(x = Voice, y = VCR)) +
  #geom_violin( color="#339966") +
  labs(x="Varianten nach Stimmhaftigkeit") +
  facet_wrap(~session, labeller = labeller(session = 
    c("_0283_TUX_jungII" = "283: TUX, jung, m",
      "_0356_PASS_jungII" = "356: PASS, jung, w",
      "_0541_GAWE_jungII" = "541: GAWE, jung, m")))+
    geom_jitter(position=position_jitter(0.2), color="#339966")+
  geom_boxplot(width = 0.2, fill = "white", color = "#9966FF")+
  theme_bw()
```

#plot relative duration as scatterplot
```{r}
ggplot(all_param, aes(x = phon.x, y = dur_rel)) +
  #geom_violin( color="#339966") +
     facet_wrap(~session, labeller = labeller(session = 
    c("_0283_TUX_jungII" = "283: TUX, jung, m",
      "_0356_PASS_jungII" = "356: PASS, jung, w",
      "_0541_GAWE_jungII" = "541: GAWE, jung, m")))+
  geom_point(color="#9966FF")+
   stat_summary(
   geom = "point",
   fun = "mean",
  col = "#339966",
   size = 3,
   shape = 24,
   fill = "#339966"
  )+
  labs(y="relative Konsonantendauer", x="Variante")+
  theme_bw()
```
#plot VCratio as scatterplot
```{r}
ggplot(all_param, aes(x = phon.x, y = V_C_ratio)) +
  #geom_violin( color="#339966") +
     facet_wrap(~session, labeller = labeller(session = 
    c("_0283_TUX_jungII" = "283: TUX, jung, m",
      "_0356_PASS_jungII" = "356: PASS, jung, w",
      "_0541_GAWE_jungII" = "541: GAWE, jung, m")))+
  geom_point(color="#9966FF")+
   stat_summary(
   geom = "point",
   fun = "mean",
  col = "#339966",
   size = 3,
   shape = 24,
   fill = "#339966"
  )+
  labs(y="VCratio", x="Variante")+
  theme_bw()
```

#plot VCR a scatterplot
```{r}
ggplot(all_param, aes(x = phon.x, y = VCR)) +
  #geom_violin( color="#339966") +
     facet_wrap(~session, labeller = labeller(session = 
    c("_0283_TUX_jungII" = "283: TUX, jung, m",
      "_0356_PASS_jungII" = "356: PASS, jung, w",
      "_0541_GAWE_jungII" = "541: GAWE, jung, m")))+
  geom_point(color="#9966FF")+
   stat_summary(
   geom = "point",
   fun = "mean",
  col = "#339966",
   size = 3,
   shape = 24,
   fill = "#339966"
  )+
  labs(y="VCR (dB)", x="Variante")+
  theme_bw()
```

#plot VCratio as boxplot with jitter
```{r}
ggplot(all_param, aes(x = phon.x, y = V_C_ratio)) +
  #geom_violin( color="#339966") +
     facet_wrap(~session, labeller = labeller(session = 
    c("_0283_TUX_jungII" = "283: TUX, jung, m",
      "_0356_PASS_jungII" = "356: PASS, jung, w",
      "_0541_GAWE_jungII" = "541: GAWE, jung, m")))+
  geom_jitter(position=position_jitter(0.2), color="#339966")+
  geom_boxplot(width = 0.2, fill = "white", color = "#9966FF")+
  labs(y="VCratio", x="Realisierung")+
  theme_bw()
```


#plot violinplot of VCVratio (not needed)
```{r}
ggplot(all_param, aes(x = phon.x, y = VCV_ratio)) +
  geom_violin( color="#339966") +
     facet_wrap(~session, labeller = labeller(session = 
    c("_0283_TUX_jungII" = "283: TUX, jung, m",
      "_0356_PASS_jungII" = "356: PASS, jung, w",
      "_0541_GAWE_jungII" = "541: GAWE, jung, m")))+
  geom_boxplot(width = 0.2, fill = "white", color = "#9966FF")+
  labs(y="VCratio", x="Realisierung")+
  theme_bw()
```


#standardisation 
```{r}
all_param$VCR_v1_st <- (all_param$VCR_v1 - mean(all_param$VCR_v1)) / sd(all_param$VCR_v1)
all_param$VCR_st <- (all_param$VCR - mean(all_param$VCR)) / sd(all_param$VCR)
all_param$V_C_ratio_st <- (all_param$V_C_ratio - mean(all_param$V_C_ratio)) / sd(all_param$V_C_ratio)
all_param$dur_rel_st <- (all_param$dur_rel - mean(all_param$dur_rel)) / sd(all_param$dur_rel)
view(all_param)
```

#plot violinplot for voice (not needed)
```{r}
ggplot(all_param, aes(x = Voice, y = VCR_v1)) +
  geom_violin( color="#339966") +
  geom_boxplot(width = 0.2, fill = "white", color = "#9966FF")+
  labs(x="Realisierung") +
  facet_wrap(~phon.x)+
  theme_bw()
```

#plot duration of preceeding vowel
```{r}
all_param %>%
ggplot(aes(x = phon.x, y = v1_dur)) +
  geom_violin( color="#339966") +
  geom_boxplot(width = 0.2, fill = "white", color = "#9966FF")+
  facet_wrap(~session)+
  labs(x="Realisierung", y= "Dauer des vorangegangenen Vokals") +
  theme_bw()
```

```{r}
sum<- all_param |> group_by(session, phon.x) |> summarise(n=n())
#view(sum)

```

#calculate rel frequencies for voice
```{r}
voice_frq <-all_param %>%
  group_by(session, phon.x) %>%   # Group by x-variable first
    mutate(total_per_Set = n()) %>%  # Total observations per x-category
    group_by(Voice, phon.x, session)%>%  # Then group by subgroup (am)
    dplyr::summarize(
      count = n(),
      freq_within_Set = count / first(total_per_Set), # freq within x-category
      .groups = "drop",
    )
show(voice_frq)
```

#plot relative frequencies for voice
```{r}
voice_frq %>%
  ggplot(aes(x= phon.x, y=freq_within_Set, fill=Voice))+
    geom_col(position="stack")+
    facet_wrap(~session, labeller = labeller(session = 
    c("_0283_TUX_jungII" = "283: TUX, jung, m",
      "_0356_PASS_jungII" = "356: PASS, jung, w",
      "_0541_GAWE_jungII" = "541: GAWE, jung, m")))+
    theme_bw()+
    scale_fill_manual(values=c("#339966", "#9966FF", "#6600FF"))+
    geom_text(aes(label=format(round(freq_within_Set, 2), nsmall = 2)), position = position_stack(0.5),check_overlap = TRUE, size = 4)+
    labs(y="relative Häufigkeit", x="Variante")

```

# scatterplot of realisierungen 
```{r}
#view(all_param)
ggplot(all_param, aes(x=VCR_st, y=dur_rel_st, color=phon.x, shape = session)) + geom_point()+
  labs(x="VCR (standardisiert)", y="relative Konsonantendauer (standardisiert)")+
  scale_color_manual(values=c("#339966", "#9966FF", "#6600FF"))+
  #geom_smooth(method = "loess",
             # formula = y ~ x, col="#9966FF")+
  theme_bw()

cor.test(all_param$VCR_st, all_param$dur_rel_st, alternative="greater", method="pearson")
cor.test(spk0283$VCR_st, spk0283$dur_rel_st, alternative="greater")
cor.test(spk0541$VCR_st, spk0541$dur_rel_st, alternative="greater")
cor.test(spk0356$VCR_st, spk0356$dur_rel_st, alternative="greater")
```


```{r}
all_param <- all_param %>% mutate(phon_num = as.numeric(phon.x))
view(all_param)
```


#filter by speaker in order to fit individual models
```{r}
spk0356 <- all_param %>%
  filter(grepl("0356", bundle))
view(spk0356)

spk0283 <- all_param %>%
  filter(grepl("0283", bundle))
#view(spk0283)

spk0541 <- all_param %>%
  filter(grepl("0541", bundle))
#view(spk0541)
```

##Erklärung Variablen: 
#VCR_v1: Vowel consonant itensity ratio mit V1
#VCR_v2: Vowel consonant itensity ratio mit V2
#V_C_ratio:duration_phon / VC_duration
#dur_rel:duration_phon / VCV_duration
#_st: standardisiert 




#test to see if categories behave ordinally
```{r}
plot.xmean.ordinaly(phon.x~VCR, data = spk0541)
```

#test with a linear model significance of v2 on VCR (not significant)
```{r}
table(all_param$v2)
all_param["v2"][all_param["v2"]=="lˈ"] <- "li"

all_param$v2<- factor(all_param$v2)
all_param <- all_param%>% filter(v2 != "V2")
m_VCR <- lm(VCR ~ v2, data = all_param)
summary(m_VCR)

```



#correlation tests (Spearman's rank correlation rho)
```{r}
cor.test(spk0283$dur_rel_st,spk0283$phon_num, method="spearman", exact = FALSE, alternative="greater")
cor.test(spk0541$dur_rel_st,spk0541$phon_num, method="spearman", exact = FALSE, alternative="greater")
cor.test(spk0356$dur_rel_st,spk0356$phon_num, method="spearman", exact = FALSE, alternative="greater")

cor.test(spk0283$VCR_st, spk0283$phon_num,method="spearman", exact = FALSE, alternative="greater")
cor.test(spk0541$VCR_st, spk0541$phon_num,method="spearman", exact = FALSE, alternative="greater")
cor.test(spk0356$VCR_st,spk0356$phon_num, method="spearman", exact = FALSE, alternative="greater")
```

# correlation test for VCratio (Spearman's rank correlation rho)
```{r}
cor.test(spk0283$V_C_ratio, spk0283$phon_num,method="spearman", exact = FALSE, alternative="less")
cor.test(spk0541$V_C_ratio, spk0541$phon_num,method="spearman", exact = FALSE, alternative="less")
cor.test(spk0356$V_C_ratio, spk0356$phon_num, method="spearman", exact = FALSE, alternative="less")
```


#models 
```{r}
m_0283 <- clm(phon.x ~ VCR_st + dur_rel_st, data = spk0283)
summary(m_0283)
m_0541 <- clm(phon.x ~ VCR_st + dur_rel_st, data = spk0541)
summary(m_0541)
m_0356 <- clm(phon.x ~ VCR_st + dur_rel_st, data = spk0356)
summary(m_0356)
```

#LRTs
```{r}
m_0283_null <- clm(phon.x ~1, data = spk0283)
m_0283_v2 <- clm(phon.x ~ dur_rel_st, data = spk0283)
m_0283_v1 <- clm(phon.x ~ VCR_st, data = spk0283)
anova(m_0283_v1, m_0283)
anova(m_0283_v2, m_0283)
anova(m_0283_null, m_0283)
```

```{r}
m_0541_null <- clm(phon.x ~1, data = spk0541)
m_0541_v2 <- clm(phon.x ~ dur_rel_st, data = spk0541)
m_0541_v1 <- clm(phon.x ~ VCR_st, data = spk0541)
anova(m_0541_v1, m_0541)
anova(m_0541_v2, m_0541)
anova(m_0541_null, m_0541)
```

```{r}
m_0356_null <- clm(phon.x ~1, data = spk0356)
m_0356_v2 <- clm(phon.x ~ dur_rel_st, data = spk0356)
m_0356_v1 <- clm(phon.x ~ VCR_st, data = spk0356)
anova(m_0356_v1, m_0356)
anova(m_0356_v2, m_0356)
anova(m_0356_null, m_0356)
```


# predicted probabilities per speaker
```{r}
emmeans(m_0541, pairwise ~VCR_st, adjust="none")
ref_grid(m_0541)
plot_model(m_0541, type="emm", terms="dur_rel [all]", title= "Prognostizierte Wahrscheinlichkeit der Varianten (GP 541)", axis.title= c("relative Konsonantendauer","Wahrscheinlichkeit der Variante"))
```

```{r}
emmeans(m_0356, pairwise ~VCR_st, adjust="none")
ref_grid(m_0356)
plot_model(m_0356, type="emm", terms="dur_rel [all]", title= "Prognostizierte Wahrscheinlichkeit der Varianten (GP 356)", axis.title= c("relative Konsonantendauer","Wahrscheinlichkeit der Variante"))
```

```{r}
emmeans(m_0283, pairwise ~VCR_st, adjust="none")
ref_grid(m_0283)
plot_model(m_0283, type="emm", terms="dur_rel [all]", title= "Prognostizierte Wahrscheinlichkeit der Varianten (GP 283)", axis.title= c("relative Konsonantendauer","Wahrscheinlichkeit der Variante"))
```


#some more tests (maybe investigate what these mean lol)
```{r}
emmeans(m_0283, ~VCR_st, adjust="none")
emmeans(m_0283, ~cut|VCR_st, mode="linear.predictor")

emtrends(m_0283, var="VCR_st", specs= ~VCR_st, infer=T)
tab_model(m_0283)
tab_model(m_0541)
tab_model(m_0356)
```

#fit models for voice
```{r}
v_0541 <- clm(Voice ~ VCR_st, data = spk0541)
summary(v_0541)

v_0356 <- clm(Voice ~ VCR_st, data = spk0356)
summary(v_0356)
v_0283 <- clm(Voice ~ VCR_st, data = spk0283)
summary(v_0283)
```

#calculate mean for VCR
```{r}
# Specify data frame
group_mean <- all_param %>%
    # Specify group indicator, column, function
    group_by(phon.x) %>%
    # Calculate the mean of the "Frequency" column for each group
    summarise_at(vars(VCR),
                 list(Mean_VCR = mean))

# Print the resulting summary data frame
print(group_mean)

```

#calculate mean for rel dur
```{r}
# Specify data frame
group_mean <- all_param %>%
    # Specify group indicator, column, function
    group_by(phon.x) %>%
    # Calculate the mean of the "Frequency" column for each group
    summarise_at(vars(dur_rel),
                 list(Mean_reldur = mean))

# Print the resulting summary data frame
print(group_mean)
```

#calculate  means and sd
```{r}
ag_vcr<- aggregate(VCR ~ phon.x, all_param, function(x) c(mean = mean(x), sd = sd(x)))
view(ag_vcr)
ag_vcr1<- aggregate(VCR ~ Voice, all_param, function(x) c(mean = mean(x), sd = sd(x)))
view(ag_vcr1)
ag_reldur<- aggregate(dur_rel ~ phon.x, all_param, function(x) c(mean = mean(x), sd = sd(x)))
view(ag_reldur)
ag_VCratio<- aggregate(V_C_ratio ~ phon.x, all_param, function(x) c(mean = mean(x), sd = sd(x)))
view(ag_VCratio)
```

#descriptive statistics for variants 
```{r}
plo <- all_param %>%
  filter(phon.x=="b")
fric <- all_param %>%
  filter(phon.x=="β")
appr <- all_param %>%
  filter(phon.x=="β̞")

summary(plo$duration_phon)
summary(fric$duration_phon)
summary(appr$duration_phon)
summary(plo$v1_dur)
summary(fric$v1_dur)
summary(appr$v1_dur)
```


#summary voice
```{r}
sum_voice<- all_param |> group_by(phon.x, Voice) |> summarise(n=n())
view(sum_voice)
```

