---
title: "analysis_auditiv"
author: "Magdalena"
date: "2026-05-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("scales")
#install.packages("lmtest")
#install.packages("gt")
library(tidyverse)
library(wrassp)
library(ggplot2)
library(scales) 
library(ordinal)
library(lme4)
library(lmtest)
library(gtsummary)
library(gt)
library(stats)
```

#load csv as df 
```{r}
#rm(list=ls(all.names=TRUE))
summary(x <- read.table(file="C:/Users/magdi/OneDrive/Dokumente/Studium/Deutsche Philologie/MASTER Deutsche Philologie/MA Arbeit/daten/Auswertung_04_vollst.csv",
header=TRUE, sep=";", quote="", comment.char="",
stringsAsFactors=TRUE))
view(x)
```

#load kontrollierte Sprachdaten
```{r}
#rm(list=ls(all.names=TRUE))
summary(x1 <- read.table(file="C:/Users/magdi/OneDrive/Dokumente/Studium/Deutsche Philologie/MASTER Deutsche Philologie/MA Arbeit/daten/kontrl_daten.csv",
header=TRUE, sep=";", quote="", comment.char="",
stringsAsFactors=TRUE))
view(x1)
```


#clean up
```{r}
clean_df <- subset(x, select = -c(X, X.1, X.2, X.3, Akustisch))
view(clean_df)
```

#calculate  sums 
```{r}
sum_df<- clean_df |> group_by(Setting) |> summarise(n=n())
sum_var<- clean_df |> group_by(Realisierung) |> summarise(n=n())
words<- clean_df |> group_by(Belegwort) |> summarise(n=n())
sum_V2<- clean_df |> group_by(V2_TAG) |> summarise(n=n())
sum_V1<- clean_df |> group_by(V1_TAG) |> summarise(n=n())
sum_dialektal<- clean_df |> group_by(V1_Dialektal, Setting) |> summarise(n=n())
sum_ort<- clean_df |> group_by(Ort) |> summarise(n=n())
sum_Lexem<- clean_df |> group_by(Lexem) |> summarise(n=n())
view(sum_Lexem)
```
#add level for V1_dialektal (so that we can add "Nein" and dont have NA values)
```{r}
levels(clean_df$V1_Dialektal) <- c(levels(clean_df$V1_Dialektal), "nein")
#view(sum_dialektal)
#droplevels(clean_df$V1_Dialektal)
```

# add columns for region, age, gender
```{r}
clean_df$Ort <- ifelse(clean_df$GP == 268|clean_df$GP == 274|clean_df$GP == 280|clean_df$GP == 283, "TUX",
                         ifelse(clean_df$GP == 366|clean_df$GP == 354|clean_df$GP == 356|clean_df$GP == 371, "PASS", "GAWE"))
clean_df["V1_Dialektal"][clean_df["V1_Dialektal"]==""] <- "nein"
clean_df$Gender<- ifelse(clean_df$GP == 268|clean_df$GP == 274|clean_df$GP == 356|clean_df$GP == 366|clean_df$GP == 544|clean_df$GP == 524, "w", "m")
clean_df$Alter<- ifelse(clean_df$GP == 283|clean_df$GP == 274|clean_df$GP == 356|clean_df$GP == 371|clean_df$GP == 544|clean_df$GP == 541, "jung", "alt")
clean_df$V1_TAG <- ifelse(clean_df$V1_TAG == "high", "geschlossen", ifelse(clean_df$V1_TAG=="low", "offen", ifelse(clean_df$V1_TAG=="mid", "mittel", "Liquid")))
clean_df$V2_TAG <- ifelse(clean_df$V2_TAG == "high", "geschlossen", ifelse(clean_df$V2_TAG=="low", "offen", ifelse(clean_df$V2_TAG=="mid", "mittel", "Liquid")))
clean_df$Alter <- factor(clean_df$Alter)
clean_df$Ort <- factor(clean_df$Ort)
clean_df$Gender <- factor(clean_df$Gender)
clean_df$GP <- factor(clean_df$GP)
clean_df$Realisierung <- factor(clean_df$Realisierung, levels=c("b" ,"β", "β̞") ,ordered = TRUE)
levels(clean_df$Realisierung )
view(clean_df)
```

#add column for lexemes 
```{r}
clean_df$Belegwort<- as.character(clean_df$Belegwort)
clean_df$Lexem <- ifelse(clean_df$Belegwort=="habe ich", "habe",
clean_df$Lexem <- ifelse(clean_df$Belegwort=="glaube ich", "glaube",
clean_df$Lexem <- ifelse(clean_df$Belegwort=="aber", "aber", 
clean_df$Lexem <- ifelse(clean_df$Belegwort=="über", "über", 
clean_df$Lexem <- ifelse(grepl("arbeit", clean_df$Belegwort,ignore.case = TRUE), "arbeit-", 
clean_df$Lexem <- ifelse(grepl("liebe", clean_df$Belegwort,ignore.case = TRUE), "lieber",
clean_df$Lexem <- ifelse(grepl("über.", clean_df$Belegwort, ignore.case = TRUE), "über-", 
clean_df$Lexem <- ifelse(grepl("selbe", clean_df$Belegwort,ignore.case = TRUE), "selber",
clean_df$Lexem <- ifelse(grepl("Hobby", clean_df$Belegwort,ignore.case = TRUE), "hobby(s)",
clean_df$Lexem <- ifelse(grepl("ober", clean_df$Belegwort, ignore.case = TRUE), "ober-",NA))))))))))
view(clean_df)
```

#view/summarise lexemes to get most frequent words 
```{r}
lexemes<- clean_df |> group_by(Lexem) |> summarise(n=n())
write.csv(lexemes,"C:/Users/magdi/OneDrive/Dokumente/Studium/Deutsche Philologie/MASTER Deutsche Philologie/MA Arbeit/daten/lexeme.csv" )
#view(lexemes)
```


#plot relative Frequencies overall
```{r}
freq_data <- clean_df %>%
  group_by(Realisierung) %>%              # Group by cylinder category
  summarize(count = n(), .groups = "drop") %>%  # Count observations per group
  mutate(
    freq = count / sum(count),   # Frequency = count / total observations
    freq_label = percent(freq)   # Optional: Format as percentage (e.g., "34%")
  )
view(freq_data)
```


##calculate rel. Freqs per Gp and Setting
```{r}
set_frq <- clean_df %>%
  group_by(GP, Setting) %>%  # Group by x-variable first
    mutate(total_per_Set = n()) %>%  # Total observations per x-category
    group_by(GP,Setting, Realisierung)%>%  # Then group by subgroup (am)
    dplyr::summarize(
      count = n(),
      freq_within_Set = count / first(total_per_Set), # freq within x-category
      .groups = "drop",
    )
set_frq$Gender<- ifelse(set_frq$GP == 268|set_frq$GP == 274|set_frq$GP == 356|set_frq$GP == 366|set_frq$GP == 544|set_frq$GP == 524, "w", "m")
set_frq$Alter<- ifelse(set_frq$GP == 283|set_frq$GP == 274|set_frq$GP == 356|set_frq$GP == 371|set_frq$GP == 544|set_frq$GP == 541, "jung", "alt")
show(set_frq)
```

##plot absolute Freq per GP and Setting
```{r}
clean_df$Realisierung<- trimws(clean_df$Realisierung)
clean_df %>%
  drop_na(Realisierung)%>%
  ggplot(aes(x= Setting))+
  geom_bar(aes(fill=Realisierung))+
  facet_wrap(~GP)+
  theme_bw()+
  labs(x="Realisierung")

```

##plot rel. frequencies per GP and Setting
```{r}
set_frq %>%
  drop_na(Realisierung)%>%
  ggplot(aes(x= Setting, y=freq_within_Set, fill=Realisierung))+
    geom_col(position="stack")+
    facet_wrap(~GP+Alter+Gender)+
    theme_bw()+
  scale_fill_manual(values=c("#339966", "#9966FF", "#6600FF"))+
    geom_text(aes(label=format(round(freq_within_Set, 2), nsmall = 2)), position = position_stack(0.5),check_overlap = TRUE, size = 2)+
    labs(y="relative Häufigkeit")

```


#calculate Frequencies per Setting
```{r}
set_frq <- clean_df %>%
 # filter(Lexem != "aber") %>%
  group_by(Setting) %>%  # Group by x-variable first
    mutate(total_per_Set = n()) %>%  # Total observations per x-category
    group_by(Setting, Realisierung)%>%  # Then group by subgroup (am)
    summarize(
      count = n(),
      freq_within_Set = count / first(total_per_Set), # freq within x-category
      .groups = "drop",
    )
show(set_frq)
```

#plot frequencies per Setting
```{r}
set_frq %>%
  drop_na(Realisierung)%>%
  ggplot(aes(x= Setting, y=freq_within_Set, fill=Realisierung))+
    geom_col(position="stack")+
  scale_fill_manual(values=c("#339966", "#9966FF", "#6600FF"))+
  theme_bw()+
   geom_text(aes(label=format(round(freq_within_Set, 2), nsmall = 2)), position = position_stack(0.5),check_overlap = TRUE, size = 5)+
    labs(y="relative Häufigkeit")


```
#Calculate frequencies per age
```{r}
age_frq <- clean_df %>%
  group_by(Alter, Setting) %>%  # Group by x-variable first
    mutate(total_per_Set = n()) %>%  # Total observations per x-category
    group_by(Alter, Setting, Realisierung)%>%  # Then group by subgroup (am)
    summarize(
      count = n(),
      freq_within_Set = count / first(total_per_Set), # freq within x-category
      .groups = "drop",
    )
show(age_frq)
```

#plot frequencies per age
```{r}
age_frq %>%
  drop_na(Realisierung)%>%
  ggplot(aes(x= Setting, y=freq_within_Set, fill=Realisierung))+
    geom_col(position="stack")+
    facet_wrap(~Alter)+
    theme_bw()+
    scale_fill_manual(values=c("#339966", "#9966FF", "#6600FF"))+
  geom_text(aes(label=format(round(freq_within_Set, 2), nsmall = 2)), position = position_stack(0.5),check_overlap = TRUE, size = 4)+
    labs(y="relative Häufigkeit")

```
#calculate frequencies per Ort
```{r}
ort_frq <- clean_df %>%
  group_by(Ort, Setting) %>%  # Group by x-variable first
    mutate(total_per_Set = n()) %>%  # Total observations per x-category
    group_by(Ort, Setting, Realisierung)%>%  # Then group by subgroup (am)
    summarize(
      count = n(),
      freq_within_Set = count / first(total_per_Set), # freq within x-category
      .groups = "drop",
    ) #%>%
  #mutate( freq_within_Set = percent(freq_within_Set))
#grouped_frq$freq_within_Set <- format(round(grouped_frq$freq_within_Set, 2), nsmall = 2)
show(ort_frq)
```


#plot frequencies per region
```{r}
ort_frq %>%
  drop_na(Realisierung)%>%
  ggplot(aes(x= Setting, y=freq_within_Set, fill=Realisierung))+
    geom_col(position="stack")+
    facet_wrap(~Ort)+
    theme_bw()+
    scale_fill_manual(values=c("#339966", "#9966FF", "#6600FF"))+
    geom_text(aes(label=format(round(freq_within_Set, 2), nsmall = 2)), position = position_stack(0.5),check_overlap = TRUE, size = 4)+
    labs(y="relative Häufigkeit")

```
#calculate frequencies per Gender
```{r}
gen_frq <- clean_df %>%
  #filter(GP !="280", GP!="283") %>%
  group_by(Gender, Setting) %>%  # Group by x-variable first
    mutate(total_per_Set = n()) %>%  # Total observations per x-category
    group_by(Gender, Setting, Realisierung)%>%  # Then group by subgroup (am)
    dplyr::summarize(
      count = n(),
      freq_within_Set = count / first(total_per_Set), # freq within x-category
      .groups = "drop",
    ) #%>%
  #mutate( freq_within_Set = percent(freq_within_Set))
#grouped_frq$freq_within_Set <- format(round(grouped_frq$freq_within_Set, 2), nsmall = 2)
show(gen_frq)

sum_gen<- clean_df |> group_by(Gender) |> summarise(n=n())
view(sum_gen)
```


#plot frequencies per Gender
```{r}
gen_frq %>%
  drop_na(Realisierung)%>%
  ggplot(aes(x= Setting, y=freq_within_Set, fill=Realisierung))+
    geom_col(position="stack")+
    facet_wrap(~Gender)+
    theme_bw()+
    scale_fill_manual(values=c("#339966", "#9966FF", "#6600FF"))+
    geom_text(aes(label=format(round(freq_within_Set, 2), nsmall = 2)), position = position_stack(0.5),check_overlap = TRUE, size = 4)+
    labs(y="relative Häufigkeit")

```

#calculate frequencies per Haupttonvokal
```{r}
ht_frq <- clean_df %>%
  group_by(V1_Dialektal) %>%  # Group by x-variable first
    mutate(total_per_Set = n()) %>%  # Total observations per x-category
    group_by(V1_Dialektal,Realisierung)%>%  # Then group by subgroup (am)
    summarize(
      count = n(),
      freq_within_Set = count / first(total_per_Set), # freq within x-category
      .groups = "drop",
    ) #%>%
  #mutate( freq_within_Set = percent(freq_within_Set))
#grouped_frq$freq_within_Set <- format(round(grouped_frq$freq_within_Set, 2), nsmall = 2)
show(ht_frq)
```
#plot frequencies per Gender
```{r}
ht_frq %>%
  drop_na(Realisierung)%>%
  ggplot(aes(x= V1_Dialektal, y=freq_within_Set, fill=Realisierung))+
    geom_col(position="stack")+
    #facet_wrap(~V1_Dialektal)+
    theme_bw()+
    scale_fill_manual(values=c("#339966", "#9966FF", "#6600FF"))+
    geom_text(aes(label=format(round(freq_within_Set, 2), nsmall = 2)), position = position_stack(0.5),check_overlap = TRUE, size = 4)+
    labs(y="relative Häufigkeit", x="dialektaler Haupttonvokal (ja/nein)")

```


#calcualte rel. frequencies for lexemes
```{r}
lex_frq <- clean_df %>%
  group_by(Lexem) %>%  # Group by x-variable first
    mutate(total_per_Lex = n()) %>%  # Total observations per x-category
    group_by(Lexem, Realisierung)%>%  # Then group by subgroup (am)
    summarize(
      count = n(),
      freq_within_Lex = count / first(total_per_Lex), # freq within x-category
      .groups = "drop",
    ) #%>%
  #mutate( freq_within_Set = percent(freq_within_Set))
#grouped_frq$freq_within_Set <- format(round(grouped_frq$freq_within_Set, 2), nsmall = 2)
view(lex_frq)
```

#plot frequencies for lexemes 
```{r}
lex_frq %>%
  drop_na(Realisierung)%>%
  #drop_na(Lexem)%>%
  ggplot(aes(x=Lexem, y=freq_within_Lex, fill=Realisierung))+
    geom_col(position="stack")+
    #facet_wrap(~Ort)+
    theme_bw()+
    scale_fill_manual(values=c("#339966", "#9966FF", "#6600FF"))+
    geom_text(aes(label=format(round(freq_within_Lex, 2), nsmall = 2)), position = position_stack(0.5),check_overlap = TRUE, size = 4)+
    labs(y="relative Häufigkeit", x="Morphem")
```

#plot frequencies for lexemes (nur über)
```{r}
lex_frq %>%
  drop_na(Realisierung)%>%
  filter %>%
  ggplot(aes(x=Lexem, y=freq_within_Lex, fill=Realisierung))+
    geom_col(position="stack")+
    #facet_wrap(~Ort)+
    theme_bw()+
    scale_fill_manual(values=c("#339966", "#9966FF", "#6600FF"))+
    geom_text(aes(label=format(round(freq_within_Lex, 2), nsmall = 2)), position = position_stack(0.5),check_overlap = TRUE, size = 4)+
    labs(y="relative Häufigkeit", x="Morphem")
```


#rel frequencies for V1 Tag
```{r}
tag_frq <- clean_df %>%
  group_by(V1_TAG) %>%  # Group by x-variable first
    mutate(total_per_TAG = n()) %>%  # Total observations per x-category
    group_by(V1_TAG, binary)%>%  # Then group by subgroup (am)
    summarize(
      count = n(),
      freq_within_TAG = count / first(total_per_TAG), # freq within x-category
      .groups = "drop",
    ) 

view(tag_frq)

tag_frq$V1_TAG <- factor(tag_frq$V1_TAG, levels=c("offen", "mittel", "geschlossen","Liquid", ordered=TRUE ))
tag_frq<- tag_frq %>% filter(binary=="NONPLOS")
show(tag_frq)
```

#plot freq for V1 Tag
```{r}
tag_frq %>%
  drop_na(binary)%>%
  drop_na(V1_TAG)%>%
  ggplot(aes(x=V1_TAG, y=freq_within_TAG, group=1))+
    geom_line() + geom_point() +theme_bw() +
    labs(y="Frequenz der spirantisierten Formen", x= "Grad der Öffnung des vorangegangenen Vokals")
```

#rel freq v1
```{r}
tag1_frq <- clean_df %>%
  group_by(V1_TAG) %>%  # Group by x-variable first
    mutate(total_per_TAG = n()) %>%  # Total observations per x-category
    group_by(V1_TAG, Realisierung)%>%  # Then group by subgroup (am)
    summarize(
      count = n(),
      freq_within_TAG = count / first(total_per_TAG), # freq within x-category
      .groups = "drop",
    ) 

#view(tag1_frq)

```
#plot v1 (bar chart)
```{r}
tag1_frq %>%
  drop_na(Realisierung)%>%
    filter(V1_TAG !="Liquid") %>%
  ggplot(aes(x= V1_TAG, y=freq_within_TAG, fill=Realisierung))+
    geom_col(position="stack")+
    #facet_wrap(~V1_Dialektal)+
    theme_bw()+
    scale_fill_manual(values=c("#339966", "#9966FF", "#6600FF"))+
    geom_text(aes(label=format(round(freq_within_TAG, 2), nsmall = 2)), position = position_stack(0.5),check_overlap = TRUE, size = 4)+
    labs(y="relative Häufigkeit", x="Öffnungsgrad des vorangegangenen Vokal")

```


#rel freq v2
```{r}
tag2_frq <- clean_df %>%
  group_by(V2_TAG) %>%  # Group by x-variable first
    mutate(total_per_TAG = n()) %>%  # Total observations per x-category
    group_by(V2_TAG, Realisierung)%>%  # Then group by subgroup (am)
    summarize(
      count = n(),
      freq_within_TAG = count / first(total_per_TAG), # freq within x-category
      .groups = "drop",
    ) 

#view(tag2_frq)

```

#plot v2 (bar chart)
```{r}
tag2_frq %>%
  drop_na(Realisierung)%>%
    filter(V2_TAG !="Liquid") %>%
  ggplot(aes(x= V2_TAG, y=freq_within_TAG, fill=Realisierung))+
    geom_col(position="stack")+
    #facet_wrap(~V1_Dialektal)+
    theme_bw()+
    scale_fill_manual(values=c("#339966", "#9966FF", "#6600FF"))+
    geom_text(aes(label=format(round(freq_within_TAG, 2), nsmall = 2)), position = position_stack(0.5),check_overlap = TRUE, size = 4)+
    labs(y="relative Häufigkeit", x="Öffnungsgrad des nachfolgenden Vokal")

```


#plot Liquide only 
```{r}
rows <- list()

# Add rows one by one
rows[[length(rows) + 1]] <- data.frame(V_TAG="vorangegangener Liquid", Realisierung= "b", Count= 20, freq_within_SET= 0.1492537 )
rows[[length(rows) + 1]] <- data.frame(V_TAG="vorangegangener Liquid", Realisierung="β", Count= 100, freq_within_SET=0.7462687 )
rows[[length(rows) + 1]]<- data.frame(V_TAG="vorangegangener Liquid", Realisierung="β̞",Count=  14,freq_within_SET= 0.1044776)

rows[[length(rows) + 1]] <-data.frame(V_TAG="nachfolgender Liquid", Realisierung="b", Count= 8, freq_within_SET= 0.33333333 )
rows[[length(rows) + 1]]<- data.frame(V_TAG="nachfolgender Liquid", Realisierung="β", Count= 15, freq_within_SET= 0.6250000 )
rows[[length(rows) + 1]]<- data.frame(V_TAG="nachfolgender Liquid", Realisierung="β̞",Count=  1, freq_within_SET= 0.041666676)

tag_frq<- do.call(rbind, rows)
view(tag_frq)

```

```{r}
tag_frq %>%
  drop_na(Realisierung)%>%
  ggplot(aes(x= V_TAG, y=freq_within_SET, fill=Realisierung))+
    geom_col(position="stack")+
    #facet_wrap(~V1_Dialektal)+
    theme_bw()+
    scale_fill_manual(values=c("#339966", "#9966FF", "#6600FF"))+
    geom_text(aes(label=format(round(freq_within_SET, 2), nsmall = 2)), position = position_stack(0.5),check_overlap = TRUE, size = 4)+
    labs(y="relative Häufigkeit", x="vorangegangener und nachfolgender Liquid")
```



#change Realisierung to binary category

```{r}
clean_df <- clean_df %>% 
  mutate(binary=ifelse(Realisierung=="b", "PLOS", "NONPLOS"))
clean_df$binary <- factor(clean_df$binary, levels=c("PLOS", "NONPLOS"))

x1 <- x1 %>% 
  mutate(binary=ifelse(Realisierung=="b", "PLOS", "NONPLOS"))
x1$binary <- factor(x1$binary, levels=c("PLOS", "NONPLOS"))
```


```{r}
x1 <- x1 %>%
   mutate(Setting = "WSUE1/LESN") 
contrl_freq <- x1 %>%
  group_by(Setting, GP) %>%  # Group by x-variable first
    mutate(total_per_Set = n()) %>%  # Total observations per x-category
    group_by(GP,Setting, binary)%>%  # Then group by subgroup (am)
    summarize(
      count = n(),
      freq_within_Set = count / first(total_per_Set), # freq within x-category
      .groups = "drop",
    )

show(contrl_freq)

```

```{r}
##calculate rel. Freqs per Gp and Setting
set_frq <- clean_df %>%
  group_by(Setting, GP) %>%  # Group by x-variable first
    mutate(total_per_Set = n()) %>%  # Total observations per x-category
    group_by(GP, Setting, binary)%>%  # Then group by subgroup (am)
    dplyr::summarize(
      count = n(),
      freq_within_Set = count / first(total_per_Set), # freq within x-category
      .groups = "drop",
    )
view(set_frq)
```


```{r}
all_sets_frq <- rbind(set_frq, contrl_freq)
view(all_sets_frq)
```


#create line plot für frequenz spirantisierte Formen for speakers individually
```{r}

all_sets_frq<- all_sets_frq %>% filter(binary=="NONPLOS")
show(all_sets_frq)

all_sets_frq$Setting<-as.character(all_sets_frq$Setting)
all_sets_frq$binary<-as.character(all_sets_frq$binary)

#view(all_sets_frq)

all_sets_frq %>%
  drop_na(binary)%>%
  drop_na(Setting)%>%
  ggplot(aes(x=Setting, y=freq_within_Set, group=GP, color=GP))+
    geom_line() + geom_point() +theme_bw() +
    labs(y="Frequenz der spirantisierten Formen", x= "Setting")

```




#create line plot für frequenz spirantisierte Formen pro Ort
```{r}
set_frq <- clean_df %>%
  group_by(Setting, Ort) %>%  # Group by x-variable first
    mutate(total_per_Set = n()) %>%  # Total observations per x-category
    group_by(Setting, Ort, binary)%>%  # Then group by subgroup (am)
    dplyr::summarize(
      count = n(),
      freq_within_Set = count / first(total_per_Set), # freq within x-category
      .groups = "drop",
    )

set_frq<- set_frq %>% filter(binary=="NONPLOS")
show(set_frq)

set_frq$Setting<-as.character(set_frq$Setting)
set_frq$binary<-as.character(set_frq$binary)
set_frq$Ort<-as.character(set_frq$Ort)


set_frq[nrow(set_frq) + 1,] = list("WSUE1/LESN","GAWE", "NONPLOS", 7, 0.65)
set_frq[nrow(set_frq) + 1,] = list("WSUE1/LESN", "PASS","NONPLOS", 12, 0.48)
set_frq[nrow(set_frq) + 1,] = list("WSUE1/LESN", "TUX","NONPLOS", 13, 0.43)


set_frq %>%
  drop_na(binary)%>%
  drop_na(Setting)%>%
  ggplot(aes(x=Setting, y=freq_within_Set, group=1))+
  facet_wrap(~Ort) +
    geom_line() + geom_point() +theme_bw() +
    labs(y="Frequenz der spirantisierten Formen", x= "Setting")

```


#generalized mixed effects model (binomial)
```{r}
# null model (for comparison)
nullmodel <- glmer(Realisierung ~ 1+(1|GP), data = clean_df, family = binomial)
summary(nullmodel)

#full model 
m01 <- glmer(Realisierung ~ V1_Dialektal + Alter+ Gender + Ort +Setting+V1_TAG +V2_TAG +(1|GP), data = clean_df, family = binomial)
summary(m01, corr = FALSE)

#without insignificant effects 
m02 <- glmer(Realisierung ~ V1_Dialektal + Ort +Setting+ V1_TAG +(1|GP), data = clean_df, family = binomial)
summary(m02, corr = FALSE)

```

#convert estimates for FEs to propabilities 
```{r}
plogis(-1.28554)
plogis(-0.41987) 
plogis(1.22923)
plogis(-0.62240)
```

#some performance measures for model
```{r}
performance(m01)
performance(m05)
```

#likelihood ratio test
#compares two models (key factors: LogLik should increase, if not model isnt a better fit, p-Value shows significance)
```{r}
l_ratio_test <- lrtest(m01, m02)
l_ratio_test <- lrtest(m05, m06)
l_ratio_test
```


```{r}
ag<- aggregate(freq_within_Set ~ Realisierung, set_frq, function(x) c(mean = mean(x), sd = sd(x)))
view(ag)
#view(set_frq)
```



