---
title: "GAWE_DB"
author: "Magdalena"
date: "2026-01-07"
output: html_document
---
##script to convert wav an corresponding TextGRid files to emu db 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries
```{r}
library(tidyverse)
library(emuR)
library(wrassp)
library(ggplot2)
```
## create source directory (where source files are stored)
```{r}
sourceDir <- "C:/Users/magdi/OneDrive/Desktop/ipsR/all"
```
##define target Directory (where to store the db)
```{r}
targetDir <- "C:/Users/magdi/OneDrive/Desktop/ipsR/emu_databases"
```


#list files
```{r}
list.files(sourceDir)
```

#Convert Collection of TextGrids to EmuDB format
```{r}
convert_TextGridCollection(sourceDir, 
                           dbName = "all",
                           targetDir = targetDir,
                           verbose = FALSE)
```
#create DB 
```{r}
DB <- load_emuDB(file.path(targetDir, "all_emuDB"),
                       verbose = FALSE)
```
#show summary of DB
```{r}
summary(DB)
```
#open DB in websocket
```{r}
serve(DB, useViewer = FALSE)
```

## Praat TextGrid files do not establish a hierarchy --> therfore requerying is not possible --> autobuilding hierarchies between each annotation layer (step by step)
## you may need to close the websocket and serve again in order for it to be shown correctly
## Pay attention to the hierarchies!! (overlapping segments need MANY_TO_MANY)

##1 word-->phon
```{r}
# add linkDefinition as one has to be present for
# the autobuild function to work
add_linkDefinition(emuDBhandle = DB,
type = "ONE_TO_MANY",
superlevelName = "word",
sublevelName = "phon")

# envoke autobuild function to build hierarchy for converted TextGridCollection
autobuild_linkFromTimes(emuDBhandle = DB,
superlevelName = "word",
sublevelName = "phon",
convertSuperlevel = FALSE)
```
##2 phon-->Phone
```{r}
# add linkDefinition as one has to be present for
# the autobuild function to work
add_linkDefinition(emuDBhandle = DB,
type = "ONE_TO_MANY",
superlevelName = "phon",
sublevelName = "dur")

# envoke autobuild function to build hierarchy for converted TextGridCollection
autobuild_linkFromTimes(emuDBhandle = DB,
superlevelName = "phon",
sublevelName = "dur",
convertSuperlevel = FALSE)
```

##3 Phone --> dur
```{r}
add_linkDefinition(emuDBhandle = DB,
type = "ONE_TO_MANY",
superlevelName = "dur",
sublevelName = "voice")

# envoke autobuild function to build hierarchy for converted TextGridCollection
autobuild_linkFromTimes(emuDBhandle = DB,
superlevelName = "dur",
sublevelName = "voice",
convertSuperlevel = FALSE)
```
