---
title: "Analysis1"
author: "Magdalena"
date: "2026-01-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Load Libraries
```{r}
library(tidyverse)
library(emuR)
library(wrassp)
library(ggplot2)
```

```{r}
targetDir <- "C:/Users/magdi/OneDrive/Desktop/ipsR/emu_databases"
```


```{r}
DB <- load_emuDB(file.path(targetDir, "all_emuDB"),
                       verbose = FALSE)
serve(DB, useViewer = FALSE)
```
# shows all labels in dur tier (i.e.V1, C, V2)
```{r}
allVC <- query(DB, "[dur =~ .*]", )
allVC <- allVC %>%filter(labels== "V1" |labels == "nV1" |labels=="C")
#view(allVC)
```

## requerying
```{r}
phon_V1_C <- requery_hier(DB, 
                    seglist = allVC, 
                    level = "phon")
phon_V1_C <- phon_V1_C %>%rename(phon=labels)
phon_V1_C <- phon_V1_C %>%mutate(uid=paste(as.character(start), bundle))
view(phon_V1_C)
```


#compute duration of V1, C
```{r}
allVC <- allVC %>%mutate(uid=paste(as.character(start), bundle))
dur_allVC <- allVC %>% mutate(duration_phon = case_when(labels == "" ~ NA, 
                                    labels != "" ~ end-start))
dur_allVC <- dur_allVC %>%rename(V_C=labels)
view(dur_allVC)
```

#compute word length
```{r}
word_length <- query(DB, query= "word=~ .*")
word_length_dur <- word_length %>% 
  mutate(duration_word = case_when(labels == "" ~ NA, 
                                    labels != "" ~ end-start))%>%
  filter(labels!= "" )%>%
  rename(word=labels)
view(word_length_dur)
view(word_length)
```

#merging tables so that every table has the duration of the respective word dur as well as phon dur

```{r}
word_phon_dur <- dur_allVC %>%
left_join(dplyr::select(word_length_dur, bundle, duration_word, word), by = "bundle")
head(word_phon_dur)
write.csv(word_phon_dur, "C:/Users/magdi/OneDrive/Desktop/ipsR/data/dur_phon.csv") ##table incl. V1 and C, phon, phon_dur
```


```{r}
word_phon_dur <- word_phon_dur %>%
  left_join(dplyr::select(phon_V1_C, phon, uid), by = "uid")
head(word_phon_dur)
```

## create table with just consonants
```{r}
allC <- word_phon_dur %>% 
  dplyr::select(V_C, phon, bundle, duration_phon, duration_word, session, uid) %>% 
  filter(V_C=="C")
allC<- na.omit(allC)
#head(allC)
#view(allC)
```
## create table with just vowels (V1) 
```{r}
allV <- word_phon_dur %>% 
  dplyr::select(V_C, word, phon, bundle, duration_phon, duration_word, uid )%>% 
  filter(V_C=="V1"| V_C=="nV1")
allV<- na.omit(allV)
#head(allV)
#view(allV)
```

#calculate VCV_duration
```{r}
allV1 <- query(DB, "[dur =V1| nV1]" )
allV1<- na.omit(allV1)
allV2 <- query(DB, "[dur =V2| nV2]")
allV2<- na.omit(allV2)
view(allV1)
view(allV2)
dur_VCV <- allC %>% mutate(VCV_duration = c((allV2$end - allV1$start)))
V1_dur<- allV  %>% mutate(phon_C=c(allC$phon))
view(V1_dur)
#head(dur_VCV)
view(dur_VCV)
```


```{r}
dur_VC <- word_phon_dur %>%
  filter(V_C=="C") 
dur_VC <- dur_VC %>%
  mutate(VC_duration = c((dur_VC$end - allV1$start)))
#view(dur_VC)
```



##calculate VCratio (V /VC) as Kleber (2018)
```{r}
V_C_ratio <- dur_VC %>%  
  mutate(V_C_ratio = c((allV$duration_phon / dur_VC$VC_duration)))%>%  
  mutate(v1_dur = c(V1_dur$duration_phon))%>%  
  mutate(VCV_ratio = c(allV$duration_phon / dur_VCV$VCV_duration))
view(V_C_ratio)
write.csv(V_C_ratio, "C:/Users/magdi/OneDrive/Desktop/ipsR/data/VC_ratio.csv", row.names=FALSE)
```

# calculate relative sound duration
```{r}
rel_dur <- dur_VCV %>%  mutate(dur_rel = c(allC$duration_phon / dur_VCV$VCV_duration))
view(rel_dur)
write.csv(rel_dur, "C:/Users/magdi/OneDrive/Desktop/ipsR/data/rel_dur.csv", row.names=FALSE)
```

#save the tables and do graphs in another script!

```{r}
allC$phon <- trimws(allC$phon)
dur_VCV$phon <- trimws(allC$phon)
rel_dur$phon <- trimws(rel_dur$phon)
V1_dur$phon_C<- trimws(V1_dur$phon_C)
```

```{r}
rel_dur <- rel_dur %>% filter(phon !="0")
ggplot(rel_dur, aes(x=phon, y=dur_rel)) + geom_boxplot()+
  labs(x="Realisierung", y= "Relative Konsonantendauer") +
  theme_bw()
```

